\propertitle

### 2019-11-15 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-15T15:18:17+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

We can approximately check how much a part of the data is irrelevant by
simply dropping it from the conditional of our probability, and see if the
latter changes appreciably. When our degree of belief is built with the
maximum-entropy method, this is done by dropping part of the constraints
and see how the resulting distribution changes. For example, we may only
retain the first two moments (means and correlations) or the first four
moments.

@@

\begin{figure}[!p]
\centering
\includegraphics[width=\linewidth]{_scripts/comparison_full.pdf}%
\\[3em]%
\includegraphics[width=\linewidth]{_scripts/trunct_comparison_marginals.pdf}%
\caption{***}
\label{fig:comparisons_marginals}
\end{figure}
% comparison_ME_sample_full_levels.nb
%
% relentropies of sample2mom, marg2mom, sample4mom, marg4mom:
% -322.72, -85.7627, -42.8005, -4.23746 nats =
% -1401.56, -372.462, -185.88, -18.403 db
% correspond to probability ratios of
% 1., 8.11718\times 10^{102}, 3.69524\times 10^{121}, 2.06704\times 10^{138}
%
% difference in relentropies between 4 mom and 2 mom:
% sample level: 279.92 nats = 1215.68 db
% marg level: 81.5252 nats =  354.059 db
% correspond to prob ratios of
% 3.69524\times 10^{121},2.5465\times 10^{35}
%
% comparison_ME_sample_full_levels.nb
%
But does the application at the full-network level lead to appreciably
different results from that at the sample level? after all we're interested
in an approximate informational sufficiency, not in an analytically exact
one. A correct answer can only be given case by case.
Figure~\ref{fig:comparisons_marginals} gives a graphical answer to this
question in the case of our first data set. The upper plot shows the
maximum-entropy distributions for a full network of $10\,000$ neurons,
constructed from two moments (\textcolor{mygreen}{\textbf{--~--~--}}) and
from four moments (\textcolor{mypurpleblue}{\textbf{------}}). The lower
plot shows the maximum-entropy distributions for the sample, from two
moments (\textcolor{mygreen}{$\bm{\triangledown}$}) and from four moments
(\textcolor{mypurpleblue}{$\bm{\square}$}), and the distributions obtained
by marginalizing the full-network distribution to the sample size, from two
moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$}) and from four
moments (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The measured
frequency distribution (\textcolor{myred}{\LARGE$\bm{\bullet}$}) is also
shown. We note the following:
\begin{itemize}
\item The application to the full-network (upper plot) leads to two
  completely different distributions (even different number of modes);
  clearly the two sets of statistics are not even approximately equivalent
  for inferential purposes.
\item The application at the sample level leads to deceivingly similar
  distributions (\textcolor{mygreen}{$\bm{\triangledown}$} and
  \textcolor{mypurpleblue}{$\bm{\square}$}), which can only be distinguished on a
  logarithmic scale. This could lead to the erroneous conclusion that the
  two statistics are approximately equivalent.
\item The difference between marginals from the full-network distribution
  and the distributions obtained at the sample level (filled vs empty
  markers) is larger than the difference between different statistics
  (triangles vs squares).
\item The marginals of the application to the full network are closer to
  the measured frequencies than the distributions obtained at the sample
  level (although this closeness is not a valid criterion for their
  goodness).
\end{itemize}
Therefore, the maximum-entropy application at the sample level and at the
full-network level lead to different results; the latter is not only more
meaningful but also superior because it shows clearly the different
informational sufficiency of the two sets of statistics.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-15T15:51:26+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

We can check how much a part of the data is irrelevant by simply dropping
it from the conditional of our probabilities, and see if the latter change
appreciably. When our probabilities are built with the maximum-entropy
method, this is done by dropping part of the constraints and see how the
resulting distribution changes. For example, we may only retain the first
two moments (means and correlations) or the first four moments. This
procedure is equivalent to calculating and comparing the probabilities of
our hypotheses about sufficiency, conditional the data, assuming their
pre-data probabilities to be equal. We perform this calculation below.

@@

\begin{figure}[!p]
\centering
\includegraphics[width=\linewidth]{_scripts/comparison_full.pdf}%
\\[3em]%
\includegraphics[width=\linewidth]{_scripts/trunct_comparison_marginals.pdf}%
\caption{***}
\label{fig:comparisons_marginals}
\end{figure}
% comparison_ME_sample_full_levels.nb
%
% relentropies of sample2mom, marg2mom, sample4mom, marg4mom:
% -322.72, -85.7627, -42.8005, -4.23746 nats =
% -1401.56, -372.462, -185.88, -18.403 db
% correspond to probability ratios of
% 1., 8.11718\times 10^{102}, 3.69524\times 10^{121}, 2.06704\times 10^{138}
%
% difference in relentropies between 4 mom and 2 mom:
% sample level: 279.92 nats = 1215.68 db
% marg level: 81.5252 nats =  354.059 db
% correspond to prob ratios of
% 3.69524\times 10^{121},2.5465\times 10^{35}
%
% comparison_ME_sample_full_levels.nb
%

 This procedure is equivalent to calculating and comparing the
probabilities of our hypotheses about sufficiency, conditional the data, if
their pre-data probability to be equal; so that for example
\begin{equation}
  \label{eq:compare_hypoth_conditional}
  \pf(\text{\small\enquote{first two moments are sufficient}} \| \text{\small data}) \propto
  \pf(\text{\small data} \|
  \text{\small\enquote{first two moments are sufficient}}).
\end{equation}

But does the application at the full-network level lead to appreciably
different results from that at the sample level? after all we're interested
in an approximate informational sufficiency, not in an analytically exact
one. A correct answer can only be given case by case.
Figure~\ref{fig:comparisons_marginals} gives a graphical answer to this
question in the case of our first data set. The upper plot shows the
maximum-entropy distributions for a full network of $10\,000$ neurons,
constructed from two moments (\textcolor{mygreen}{\textbf{--~--~--}}) and
from four moments (\textcolor{mypurpleblue}{\textbf{------}}). The lower
plot shows the maximum-entropy distributions for the sample, from two
moments (\textcolor{mygreen}{$\bm{\triangledown}$}) and from four moments
(\textcolor{mypurpleblue}{$\bm{\square}$}), and the distributions obtained
by marginalizing the full-network distribution to the sample size, from two
moments (\textcolor{mygreen}{$\bm{\blacktriangledown}$}) and from four
moments (\textcolor{mypurpleblue}{$\bm{\blacksquare}$}). The measured
frequency distribution (\textcolor{myred}{\LARGE$\bm{\bullet}$}) is also
shown. We note the following:
\begin{itemize}
\item The application to the full-network (upper plot) leads to two
  completely different distributions (even different number of modes);
  clearly the two sets of statistics are not even approximately equivalent
  for inferential purposes.
\item The application at the sample level leads to deceivingly similar
  distributions (\textcolor{mygreen}{$\bm{\triangledown}$} and
  \textcolor{mypurpleblue}{$\bm{\square}$}), which can only be distinguished on a
  logarithmic scale. This could lead to the erroneous conclusion that the
  two statistics are approximately equivalent.
\item The difference between marginals from the full-network distribution
  and the distributions obtained at the sample level (filled vs empty
  markers) is larger than the difference between different statistics
  (triangles vs squares).
\item The marginals of the application to the full network are closer to
  the measured frequencies than the distributions obtained at the sample
  level (although this closeness is not a valid criterion for their
  goodness).
\end{itemize}
Therefore, the maximum-entropy application at the sample level and at the
full-network level lead to different results; the latter is not only more
meaningful but also superior because it shows clearly the different
informational sufficiency of the two sets of statistics.
